---
title: "Cortisol Dataset Analysis"
output: html_document
---

In this problem, we consider a blind study where all the pigs are randomly assigned with a certain diet with or without chips, and we are interested in the effect of chips (so the covariate `group`) on urine cortisol. As the causal influence among hormones remains unknown, we only consider `urine cortisol` as the response variable in our analysis. We first import and clean the data as follows:

```{r}
# Import data
cortisol_data = read.csv("data/cortisol.csv", row.names=1)

# Create data frame
dat = data.frame(group = cortisol_data$group,
                 gender = cortisol_data$gender,
                 urine_cortisol = cortisol_data$Urine.Cortisol..pg.mg.,
                 serum_cortisol = cortisol_data$Serum.Cortisol..ng.ml.,
                 ACTH = cortisol_data$Serum.ACTH..pg.ml.,
                 CRH = cortisol_data$Serum.CRH..pg.ml.,
                 testosterone = cortisol_data$Testosterone..ng.ml.,
                 LH = cortisol_data$LH..ng.ml.)

head(dat[,1:3])
```

# Simple analysis

We start our analysis by visualizing the data to grasp an idea of how the data is like. 

```{r, fig.height=5, fig.width=6}
gg_color_hue = function(n, alpha = 1) {
  hues = seq(15, 375, length = n + 1)
  hcl(h = hues, l = 65, c = 100, alpha = alpha)[1:n]
}

cols = gg_color_hue(2)
cols_trans = gg_color_hue(2, alpha = 0.15)

boxplot(urine_cortisol ~ group, data = dat, col = cols, ylab = "Urine Cortisol")
```

As we can see, the groups of `Chips` and `Non-chips` show different mean, different variance with some outliers. Therefore, it is not suitable to be analyzed by any of the methods we have discussed so far. Let's try a log transformation on the response variable and see if we can get something better.

```{r}
boxplot(log(urine_cortisol) ~ group, data = dat, col = cols, ylab = "log(Urine Cortisol)")
```

As we can see, after the log transformation, the two groups show different mean but relatively similar variance with no more outliers. The transformed data can then be analyzed with a Welch's t-test with hypotheses $H_0: \mu_{C} = \mu_{NC}$ and $H_a: \mu_{C} > \mu_{NC}$. The R output is as follows:

```{r}
t.test(log(urine_cortisol) ~ group, data = dat, col = cols, alternative = "greater")
```

Since the p-value is smaller than 5%, we can conclude that the chips diet significantly increases the level of urine cortisol **on the log scale**, at the significance level of 5%. However, what can we say on the original scale of urine cortisol? This will need to involve a more complicated study with the Generalized Linear Models (GLM). 

# More complex analysis with GLM

Consider the following model where we assume that the level of urine cortisol ($Y_i$) follows a gamma distribution 

$$Y_i \sim \text{Gamma}(\mu_i, \sigma),$$

where $\mu_i$, the conditional mean of $Y_i$, is modelled by

$$\mu_i = \exp(\mathbf{x}_i^T \boldsymbol{\beta}),$$
and $\sigma$ is related to the conditional variance of $Y_i$ as $\text{var}(Y_i) = \mu_i^2 \sigma^2$. You can check out more details about the reparametrization of gamma distribution we use by loading the `gamlss` R package and checking the documentation of the `GA` function. 

```{r, warning=FALSE, message=FALSE}
library(gamlss)
```

Now let's try to fit the considered model with the `gamlss` function. 

```{r}
fit_ga_urine_cortisol = gamlss(urine_cortisol ~ group, data = dat, 
                               sigma.formula =~ 1, family = GA)
summary(fit_ga_urine_cortisol)
```

Before we draw any conlusion from the fitted model, we should first check if the model is suitable given the data. It is possible to compute standardized residuals as follows:

$$r_i = \frac{Y_i - \widehat{\mu}_i}{\widehat{\sigma}_i},$$
where $\widehat{\mu}_i$ and $\widehat{\sigma}_i$ correspond to the predicted (by the estimated model) mean and standard deviation for observation $i$. Therefore, we can perform the following model check:

```{r, fig.align='center', fig.height=5, fig.width=6}
plot(fit_ga_urine_cortisol)
```

We can also obtain the following fitted densities as follows:

```{r, fig.align='center', fig.height=5, fig.width=6}
mu_coef = fit_ga_urine_cortisol$mu.coefficients
sigma_coef = fit_ga_urine_cortisol$sigma.coefficients

mu_chips = mu_coef["(Intercept)"] 
mu_non_chips = mu_coef["(Intercept)"] + mu_coef["groupNC"] 

xx = seq(from = 0.0001, to = 10^4, length.out = 10^4)
yy_chips = dGA(xx, mu = exp(mu_chips), sigma = exp(sigma_coef))
yy_non_chips = dGA(xx, mu = exp(mu_non_chips), sigma = exp(sigma_coef))

plot(NA, xlim = c(0, 10^4), ylim = c(0, max(na.omit(yy_non_chips))), xlab = "Urine Cortisol", ylab = "PDF")

polygon(c(xx, rev(xx)), c(rep(0, length(xx)), rev(yy_chips)), border = NA, col = cols_trans[1])
lines(xx, yy_chips, col = cols[1], lwd = 1)

polygon(c(xx, rev(xx)), c(rep(0, length(xx)), rev(yy_non_chips)), border = NA, col = cols_trans[2])
lines(xx, yy_non_chips, col = cols[2], lwd = 1)

legend("topright", c("Chips", "Non-chips"), col = cols, lwd = 1)
```

```{r, fig.align='center', fig.height=5, fig.width=6}
plot(NA, xlim = c(0, 10^4), ylim = c(0, max(na.omit(yy_non_chips))), xlab = "Urine Cortisol", ylab = "PDF")

hist(dat$urine_cortisol[dat$group == "NC"], probability = TRUE, add = TRUE, col = "grey80", breaks = 3, border = cols[2])

polygon(c(xx, rev(xx)), c(rep(0, length(xx)), rev(yy_chips)), border = NA, col = cols_trans[1])
lines(xx, yy_chips, col = cols[1], lwd = 2)

polygon(c(xx, rev(xx)), c(rep(0, length(xx)), rev(yy_non_chips)), border = NA, col = cols_trans[2])
lines(xx, yy_non_chips, col = cols[2], lwd = 2)

legend("topright", c("Chips", "Non-chips"), col = cols, lwd = 1)
```

```{r, fig.align='center', fig.height=5, fig.width=6}
plot(NA, xlim = c(0, 10^4), ylim = c(0, max(na.omit(yy_non_chips))), xlab = "Urine Cortisol", ylab = "PDF")

hist(dat$urine_cortisol[dat$group == "C"], probability = TRUE, add = TRUE, col = "grey80", breaks = 10, border = cols[1])

polygon(c(xx, rev(xx)), c(rep(0, length(xx)), rev(yy_chips)), border = NA, col = cols_trans[1])
lines(xx, yy_chips, col = cols[1], lwd = 2)

polygon(c(xx, rev(xx)), c(rep(0, length(xx)), rev(yy_non_chips)), border = NA, col = cols_trans[2])
lines(xx, yy_non_chips, col = cols[2], lwd = 2)

legend("topright", c("Chips", "Non-chips"), col = cols, lwd = 1)
```

The above graphs show that the model fits the data very well. Therefore, it is reliable to draw conclusion based on this fitted gamma GLM. Since the p-value related to the `groupNC` covariate is $5.43 \times 10^{-8}\%$, and that the estimated coefficient is $-1.3016$, the p-value that corresponds to the hypotheses $H_0: \mu_{C} = \mu_{NC}$ and $H_a: \mu_{C} > \mu_{NC}$ is therefore $5.43 \times 10^{-8}\%/2 \approx 2.72\times 10^{-8}\%$. So we can conclude that the chips diet significantly increases the level of urine cortisol at the significance level of 5%. 

# Supplementary analysis: Gender

As the experiment is a blind study, we expect not to see significant difference between different genders. To verify our expectation, let's fit the model again by adding the covariate `gender`. 

```{r}
fit_ga_urine_cortisol2 = gamlss(urine_cortisol ~ group*gender, data = dat, sigma.formula =~ 1, family = GA)
summary(fit_ga_urine_cortisol2)
```

```{r, fig.align='center', fig.height=5, fig.width=6}
plot(fit_ga_urine_cortisol)
```

As we can see, the considered model fits well the data, and the covariates related to `gender` are both not significant, which verifies our expectation and concludes the analysis. 

