---
title: "Exercises on Generalized Linear Models"
output: html_document
---

# Problem 1: Students with an Infectious Disease

Suppose we are interested in the number of high school students diagnosed with an infectious disease as a function of the number of days from the initial outbreak. The data can be loaded as follows:

```{r}
students = read.csv("data/students.csv")
head(students)
```

Can you find a suitable model for this dataset?


# Problem 2: Pima Indians Diabetes

Consider a dataset on 768 women of at least 21 years old of the Pima Indian heritage. This dataset includes the following variables:

- `pregnant`: Number of times pregnant
- `glucose`: Plasma glucose concentration in an oral glucose tolerance test
- `pressure`: Diastolic blood pressure (mm Hg)
- `triceps`: Triceps skin fold thickness (mm)
- `insulin`: 2-Hour serum insulin (mu U/ml)
- `mass`: Body mass index (weight in kg/(height in m)^2)
- `pedigree`: Diabetes pedigree function
- `age`: Age of the patients (years)
- `diabetes`: Class variable (test for diabetes)

The dataset is stored in the `mlbench` R package and you can load it as follows:

```{r}
library(mlbench)
data(PimaIndiansDiabetes)
dim(PimaIndiansDiabetes)
head(PimaIndiansDiabetes)
```

Can you find a suitable model to predict the possibility of a woman (Pima Indian heritage) being diagnosed with diabetes based on these variables (or a subset of them)?



# Problem 3: Breast Cancer

Let's consider a breast cancer dataset from the `mlbench` R package. In this example, the objective is to predict whether a cancer is malignant or benign from biopsy details. This dataset includes 699 observations on 11 variables. You can load the data as follows:

```{r}
data(BreastCancer)
dim(BreastCancer)
head(BreastCancer)
table(BreastCancer$Class)
```

More information about this dataset can be found by typing `?BreastCancer` in R. Can you find a suitable model to identify benign or malignant classes based on these variables (or a subset of them)?
